---
title: "World Series Simulation Method"
author: "Alejandro Aragon"
date: "1/20/2022"
output: html_document
---

```{r, echo=FALSE}
datapath <- "F:/Data Analysis/Personal/MLBteams.csv"
```

Data pulled from [Baseball-Reference.com](https://www.baseball-reference.com)

```{r}
mlb <- read.csv(datapath, header = TRUE)
```

```{r}
#Win Rate
mlb$WR <- mlb$W/(mlb$W + mlb$L)
```

getmode() function taken from [tutorial point](https://www.tutorialspoint.com/r/r_mean_median_mode.htm)

```{r}

getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
```

How this method works:

* Two teams and associated years are passed through the function and put into a vector "teamsPlaying".
  + "nsim" was added to change how many times samples will be taken to make the results easier to read here. It was previously fixed at 100001.
* Using the teams associated years as indexes the win rate is pulled from the data frame and put in a probability vector "p" that is used to weight the sampling probability.
* `nsim` samples of the two teams are taken and put into a data frame called seriesResults.
  + This is repeated 7 times to mimic a series of 7 games.
* Using the [getmode() function](https://www.tutorialspoint.com/r/r_mean_median_mode.htm), the team that appears the most in the each of the column vectors is declared the winner of that "game" and assigned to the seriesResults vector.
* The team that returned the most frequently in the seriesResults vector is declared the winner of the simulated World Series ("WSwinner")

```{r}


WS.simulation <- function(team1, year1, team2, year2, nsim){
  set.seed(Sys.time())
  
   seriesResults <- vector()
   teamsPlaying <- c(paste(year1, team1, sep = "_"), paste(year2, team2, sep = "_"))
   
 # Pulling win rate from dataset
   team1WR <- mlb$WR[which(mlb$team == team1 & mlb$Year == year1)] 
   team2WR <- mlb$WR[which(mlb$team == team2 & mlb$Year == year2)]
   
   # List to hold the sampling results for each "game"
   gameNames <- c("Game1", "Game2", "Game3", "Game4", "Game5", "Game6", "Game7") 
   games <- list()
 
   # creating a vector of win rate as a psuedo probability
   p <- c(team1WR, team2WR)
   
   samplingResults<- replicate(7, sample(teamsPlaying, nsim, replace = TRUE, prob = p))
   
   for (i in 1:7) {
     
     # Putting most sampled team into results vector 
     seriesResults[i] <- getmode(samplingResults[,i])
     
     # Putting sampling results into previously made lists
     games[[i]] <- data.frame(Year = c(year1, year2)
                              , teams = c(team1, team2)
                              , nSampled = tabulate(match(samplingResults[,i], teamsPlaying))
                              , perSampled = c(tabulate(match(samplingResults[,i], teamsPlaying))[1]/nsim
                                               , tabulate(match(samplingResults[,i], teamsPlaying))[2]/nsim)
                              , winRate = p)
   }
   
   names(games) <- gameNames
    
    sim <- list(WSwinner = getmode(seriesResults)
                , gameResults = seriesResults
                , winRate = p
                , gameStats = games
                , samples = samplingResults
      )
    
    class(sim) <- "WS.sim"
    return(sim)
}
```

```{r}
LAD21.SFG21 <- WS.simulation("LAD", 2021,"SFG", 2021,  100001)
LAD21.SFG21
```


```{r}
WS.BOS.18.04 <- WS.simulation("BOS", 1918, "BOS", 2004, 100001)
```

```{r}
WS.BOS.18.04
```

```{r}
KCR15.CHC16 <- WS.simulation("CHC", 2016,"KCR", 2015,  100001)
KCR15.CHC16$WSwinner
KCR15.CHC16$gameResults
```

```{r}
KCR15.CHC16
```
































